---
title: "TBI"
subtitle: "SMM068 (2025-26) Group Coursework 02 - Group 03"
author:
  - "Benjamin Evans"
  - "Basmah Khan"
  - "Bong Su Kil"
  - "Ardi Wira Sudarmo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  bookdown::html_document2:
    self_contained: true
    css: style.css
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    code_folding: hide
    theme: united
    highlight: tango
    df_print: paged
  # runtime: shiny
  bookdown::pdf_document2:
    includes:
      in_header: preamble.tex
    pandoc_args:
      - "--lua-filter=answer.lua"
    number_sections: true
    toc: true
    latex_engine: xelatex
    citation_package: natbib
    keep_tex: true
# fontsize: 10pt
bibliography: references.bib
---

```{=latex}
\newpage
```

```{=html}
<h1>Information</h1>

This R Markdown document was created as part of a group assignment for SMM068 at Bayes Business School, City St George's, University of London in Term 2 2025-26.

```

```{r asthetic-header, include=FALSE, echo=FALSE}
# ============================================================================ #
# Key Information ====
# SMM068 Financial Economics (Subject CM2)
# Group Coursework 2025-26
# Group:        Group 08
# Authors (in alphabetical order):
#   - Benjamin Evans
#   - Basmah Khan
#   - Bong Su Kil
#   - Ardi Wira Sudarmo
# Professor:    Professor Ioannis Kyriakou
# Institution:  Bayes Business School - City St George's, University of London
# Date:         30 Mar 2026
#
# Description:  Term 2 group project for SMM068 Financial Economics
#               (50% of coursework grade - 10% of module grade). The R code
#               below has been exported directly from an R Markdown (.rmd) file.
#               Hence the knitr settings.
#
# Dependencies:
#   - quantmod: for downloading
#   - tidyverse: loads dplyr, tidyr, ggplot2, purrr (& others)
#   - TBI: loads dplyr, tidyr, ggplot2, purrr (& others)
#
# Acknowledgements
#   - Bodie et al., 2024, chapters 5-8 were used as a primary source for the
#     concepts coded here (in addition to the lecture notes)
# ============================================================================ #
```

```{r setup-knitr, include=FALSE, echo=FALSE}
# Setup & configuration ====
#----------------------- Initial setup (knitr settings) -----------------------#
dir.create("fig", showWarnings = FALSE)

# Defaults common to all outputs
knitr::opts_chunk$set(
  echo     = TRUE,
  message  = FALSE,
  warning  = FALSE,
  fig.align = "center",
  out.width = "100%",
  fig.path  = "fig/",
  dpi       = 300
)

# Output-specific settings
if (knitr::is_latex_output()) {
  knitr::opts_chunk$set(
    fig.width = 6,
    fig.height = 4,
    dev = "pdf",
    fig.pos = "!ht",
    out.extra = ""
  )
} else {
  knitr::opts_chunk$set(
    fig.width = 6,
    fig.height = 4,
    dev = "svglite"  # or "png"
  )
}
```

```{r setup-qol, include=FALSE, echo=FALSE}
#----------------------------- Clean environment ------------------------------#
rm(list = ls()) # Remove all objects
graphics.off() # Close all graphical devices
cat("\014") # Clean console
```

```{r load-dependencies, include=FALSE, echo=FALSE}
#------------------- Load dependencies / external libraries -------------------#
# --- Data Fetching ---
# TBI

# --- Data Manipulation & Plotting ---
library(tidyverse) # loads dplyr, tidyr, ggplot2, purrr (& others)
library(patchwork) # for combining multiple plots into one figure

# --- Reporting & Tables ---
library(kableExtra) # for general tables
library(DT) # for HTML interactive data tables
library(plotly) # for interactive html plots
library(pander) # for writing to pandoc
```

```{r html-app, echo=FALSE, results='asis', eval=knitr::is_html_output(), purl=FALSE}
#------------------------------ HTML link to app ------------------------------#
library(htmltools)

div(style = "background-color: #f8f9fa; padding: 20px; border: 1px solid #e9ecef; border-radius: 5px; text-align: center; margin-bottom: 30px;",
  h3("Bonus Interactive Dashboard Available"),
  p("!!TBI."),
  a(href = "https://3enji.shinyapps.io/smm068-a01-g08-dashboard/",
    target = "_blank",
    class = "btn btn-primary",
    style = "background-color: #007bff; color: white; padding: 10px 20px; text-decoration: none; border-radius: 5px; font-weight: bold;",
    "Launch Interactive Dashboard")
)
```

```{r custom-functions, include=FALSE, echo=FALSE}
#---------------------------- Custom QOL functions ----------------------------#
#####################################
# function: banner comments (used to to section up code)
# Usage: banner_comment("Element 1: data cleaning") -> then ctrl + v (or cmd+v)
#####################################
banner_comment <- function(text, width = 80, border = "#", fill = "-") {
  txt <- paste0(" ", text, " ")
  inner_width <- width - 2 * nchar(border)
  banner_string <- ""

  if (inner_width <= nchar(txt)) {
    banner_string <- paste0(border, txt, border)
  } else {
    pad_total <- inner_width - nchar(txt)
    pad_left <- pad_total %/% 2
    pad_right <- pad_total - pad_left

    banner_string <- paste0(
      border,
      strrep(fill, pad_left),
      txt,
      strrep(fill, pad_right),
      border
    )
  }

  cat(banner_string, "\n")
  # copy banner to allow direct pasting (requires clipr)
  clipr::write_clip(banner_string)
  # avoid [1] when printing if want to manually copy
  invisible(banner_string)
}
#####################################
# function: format p-values for text
# Usage (in-line): `r format_p_vals(ad_test_result$p.value)`
# Usage (console): format_p_vals(ad_test_result$p.value)
#####################################
format_p_vals <- function(p) {
  if (length(p) != 1L || is.na(p)) {
    stop("Error! p must be a single non-missing value")
  }
  if (p > 1) {
    stop("Error! Value greater than 1")
  }
  if (p < 0) {
    stop("Error! Value less than 0")
  }

  if (p >= 0.01) {
    paste0("= ", formatC(p, format = "f", digits = 2))
  } else if (p >= 0.001) {
    paste0("= ", formatC(p, format = "f", digits = 3))
  } else {
    "< 0.001"
  }
}
#####################################
# function: format confidence intervals for tables & text
# Usage (in-line): `r format_interval(el2_ci_normal_95[1], el2_ci_normal_95[2])`
# Usage (console): format_interval(el2_ci_normal_95[1], el2_ci_normal_95[2])
#####################################
format_interval <- function(lower, upper, digits = 3, small_interval = 5L) {
  paste0(
    "[",
    formatC(
      lower,
      format = "f",
      digits = digits,
      small.interval = small_interval,
      small.mark = " "
    ),
    ", ",
    formatC(
      upper,
      format = "f",
      digits = digits,
      small.interval = small_interval,
      small.mark = " "
    ),
    "]"
  )
}
#####################################
# function: format truncated ellipses
# Usage (in-line): `r tbi`
# Usage (console): tbi
#####################################
fmt_trunc_ellip <- function(
  x,
  digits = 7,
  ellip = "...",
  tol = 1e-12,
  trim_zeros_if_exact = TRUE
) {
  out <- rep(NA_character_, length(x))
  ok <- is.finite(x)
  scale <- 10^digits
  xt <- trunc(x[ok] * scale) / scale
  s <- formatC(xt, format = "f", digits = digits)
  add <- abs(x[ok] - xt) > tol * pmax(1, abs(x[ok]))
  if (trim_zeros_if_exact) {
    s_trim <- sub("0+$", "", s) # drop trailing zeros
    s_trim <- sub("\\.$", "", s_trim) # drop trailing decimal point
  } else {
    s_trim <- s
  }
  out[ok] <- ifelse(add, paste0(s, ellip), s_trim)
  out[is.na(x)] <- NA_character_
  out
}
#####################################
# function: function to format 6dp numbers with a space after the 3rd decimal
# digit
# Usage (in-line): TBI
# Usage (console): TBI
#####################################
format_spaced_6dp <- function(x) {
  # Format to 6 decimal places
  s <- sprintf("%.6f", x)
  # Insert space between the first 3 and last 3 decimal digits
  sub("(\\.[0-9]{3})([0-9]{3})$", "\\1 \\2", s)
}
```

# Binomial tree {#qone}

# Binomial tree figure {#qtwo}

# European vanilla put option {#qthr}

# Bermudan put option {#qfou}

# Commenting on the relative prices of {#qfiv}

# Explain how Bermudan option would change with... {#qsix}

```{=latex}
\newpage
```

<!-- # Conclusions

TBI -->

<!-- ```{=html} -->
<!-- # References -->
<!-- <div id="refs"></div> -->
<!-- ``` -->

<!-- ```{=latex} -->
<!-- \bibliography{references.bib} -->
<!-- ``` -->

```{=latex}
\newpage
\section*{Appendices}
\addcontentsline{toc}{section}{Appendices}

% --- restore default section numbering in the appendix ---
\renewcommand{\thesection}{\arabic{section}}
\renewcommand{\thesubsection}{\thesection.\arabic{subsection}}
\renewcommand{\thesubsubsection}{\thesubsection.\arabic{subsubsection}}
```

# (APPENDIX) Appendices {-}

# Supplemental methodology {#apd-supplementary-methods}

## Spreadsheet implementation (Excel) {#app:excel-impl}

This subsection outlines the structure of the accompanying Excel workbook, using
screenshots to demonstrate the layout and contents of each sheet.

... TBI

<!-- [^utility-theory]: -->
<!--     Described in more detail in the appendix (\S\@ref(apd-utility-theory)). -->

<!-- [^utility-skew]: -->
<!--     This is demonstrated in the appendix (\S\@ref(apd-utility-theory) and Table -->
<!--     \@ref(tab:exact-utility-table)) where we show that a constant relative risk -->
<!--     aversion (CRRA) utility function with $\gamma > 0$ mathematically confirms -->
<!--     that for risk-averse investors asset 1 results in a lower expected utility -->
<!--     that asset 2. --> -->
