---
title:
  "Binomial Tree Calibration and European & Bermudan Put Option Valuation for
  Microsoft Corporation (MSFT) Stock"
subtitle: "SMM068 (2025-26) Group Coursework 02 - Group 03"
author:
  - "Benjamin Evans"
  - "Basmah Khan"
  - "Bong Su Kil"
  - "Ardi Wira Sudarmo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  bookdown::html_document2:
    self_contained: true
    css: style.css
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    code_folding: hide
    theme: united
    highlight: tango
    df_print: paged
  # runtime: shiny
  bookdown::pdf_document2:
    includes:
      in_header: preamble.tex
    pandoc_args:
      - "--lua-filter=answer.lua"
    number_sections: true
    toc: true
    latex_engine: xelatex
    citation_package: natbib
    keep_tex: true
# fontsize: 10pt
bibliography: references.bib
---

```{=latex}
\newpage
```

```{=html}
<h1>Information</h1>

This R Markdown document was created as part of a group assignment for SMM068 at Bayes Business School, City St George's, University of London in Term 2 2025-26.

```

```{r asthetic-header, include=FALSE, echo=FALSE}
# ============================================================================ #
# Key Information ====
# SMM068 Financial Economics (Subject CM2)
# Group Coursework 2025-26
# Group:        Group 08
# Authors (in alphabetical order):
#   - Benjamin Evans
#   - Basmah Khan
#   - Bong Su Kil
#   - Ardi Wira Sudarmo
# Professor:    Professor Ioannis Kyriakou
# Institution:  Bayes Business School - City St George's, University of London
# Date:         30 Mar 2026
#
# Description:  Term 2 group project for SMM068 Financial Economics
#               (50% of coursework grade - 10% of module grade). The R code
#               below has been exported directly from an R Markdown (.rmd) file.
#               Hence the knitr settings.
#
# Dependencies:
#   - quantmod: for downloading
#   - tidyverse: loads dplyr, tidyr, ggplot2, purrr (& others)
#   - TBI: loads dplyr, tidyr, ggplot2, purrr (& others)
#
# Acknowledgements
#   - Bodie et al., 2024, chapters 5-8 were used as a primary source for the
#     concepts coded here (in addition to the lecture notes)
# ============================================================================ #
```

```{r setup-knitr, include=FALSE, echo=FALSE}
# Setup & configuration ====
#----------------------- Initial setup (knitr settings) -----------------------#
dir.create("fig", showWarnings = FALSE)

# Defaults common to all outputs
knitr::opts_chunk$set(
  echo     = TRUE,
  message  = FALSE,
  warning  = FALSE,
  fig.align = "center",
  out.width = "100%",
  fig.path  = "fig/",
  dpi       = 300
)

# Output-specific settings
if (knitr::is_latex_output()) {
  knitr::opts_chunk$set(
    fig.width = 6,
    fig.height = 4,
    dev = "pdf",
    fig.pos = "!ht",
    out.extra = ""
  )
} else {
  knitr::opts_chunk$set(
    fig.width = 6,
    fig.height = 4,
    dev = "svglite"  # or "png"
  )
}
```

```{r setup-qol, include=FALSE, echo=FALSE}
#----------------------------- Clean environment ------------------------------#
rm(list = ls()) # Remove all objects
graphics.off() # Close all graphical devices
cat("\014") # Clean console
```

```{r load-dependencies, include=FALSE, echo=FALSE}
#------------------- Load dependencies / external libraries -------------------#
# --- Data Fetching ---
# TBI

# --- Data Manipulation & Plotting ---
library(tidyverse) # loads dplyr, tidyr, ggplot2, purrr (& others)
library(patchwork) # for combining multiple plots into one figure

# --- Reporting & Tables ---
library(kableExtra) # for general tables
library(DT) # for HTML interactive data tables
library(plotly) # for interactive html plots
library(pander) # for writing to pandoc
```

```{r html-app, echo=FALSE, results='asis', eval=knitr::is_html_output(), purl=FALSE}
#------------------------------ HTML link to app ------------------------------#
library(htmltools)

div(style = "background-color: #f8f9fa; padding: 20px; border: 1px solid #e9ecef; border-radius: 5px; text-align: center; margin-bottom: 30px;",
  h3("Bonus Interactive Dashboard Available"),
  p("!!TBI."),
  a(href = "https://3enji.shinyapps.io/smm068-a01-g08-dashboard/",
    target = "_blank",
    class = "btn btn-primary",
    style = "background-color: #007bff; color: white; padding: 10px 20px; text-decoration: none; border-radius: 5px; font-weight: bold;",
    "Launch Interactive Dashboard")
)
```

```{r custom-functions, include=FALSE, echo=FALSE}
#---------------------------- Custom QOL functions ----------------------------#
#####################################
# function: banner comments (used to to section up code)
# Usage: banner_comment("Element 1: data cleaning") -> then ctrl + v (or cmd+v)
#####################################
banner_comment <- function(text, width = 80, border = "#", fill = "-") {
  txt <- paste0(" ", text, " ")
  inner_width <- width - 2 * nchar(border)
  banner_string <- ""

  if (inner_width <= nchar(txt)) {
    banner_string <- paste0(border, txt, border)
  } else {
    pad_total <- inner_width - nchar(txt)
    pad_left <- pad_total %/% 2
    pad_right <- pad_total - pad_left

    banner_string <- paste0(
      border,
      strrep(fill, pad_left),
      txt,
      strrep(fill, pad_right),
      border
    )
  }

  cat(banner_string, "\n")
  # copy banner to allow direct pasting (requires clipr)
  clipr::write_clip(banner_string)
  # avoid [1] when printing if want to manually copy
  invisible(banner_string)
}
#####################################
# function: format p-values for text
# Usage (in-line): `r format_p_vals(ad_test_result$p.value)`
# Usage (console): format_p_vals(ad_test_result$p.value)
#####################################
format_p_vals <- function(p) {
  if (length(p) != 1L || is.na(p)) {
    stop("Error! p must be a single non-missing value")
  }
  if (p > 1) {
    stop("Error! Value greater than 1")
  }
  if (p < 0) {
    stop("Error! Value less than 0")
  }

  if (p >= 0.01) {
    paste0("= ", formatC(p, format = "f", digits = 2))
  } else if (p >= 0.001) {
    paste0("= ", formatC(p, format = "f", digits = 3))
  } else {
    "< 0.001"
  }
}
#####################################
# function: format confidence intervals for tables & text
# Usage (in-line): `r format_interval(el2_ci_normal_95[1], el2_ci_normal_95[2])`
# Usage (console): format_interval(el2_ci_normal_95[1], el2_ci_normal_95[2])
#####################################
format_interval <- function(lower, upper, digits = 3, small_interval = 5L) {
  paste0(
    "[",
    formatC(
      lower,
      format = "f",
      digits = digits,
      small.interval = small_interval,
      small.mark = " "
    ),
    ", ",
    formatC(
      upper,
      format = "f",
      digits = digits,
      small.interval = small_interval,
      small.mark = " "
    ),
    "]"
  )
}
#####################################
# function: format truncated ellipses
# Usage (in-line): `r tbi`
# Usage (console): tbi
#####################################
fmt_trunc_ellip <- function(
  x,
  digits = 7,
  ellip = "...",
  tol = 1e-12,
  trim_zeros_if_exact = TRUE
) {
  out <- rep(NA_character_, length(x))
  ok <- is.finite(x)
  scale <- 10^digits
  xt <- trunc(x[ok] * scale) / scale
  s <- formatC(xt, format = "f", digits = digits)
  add <- abs(x[ok] - xt) > tol * pmax(1, abs(x[ok]))
  if (trim_zeros_if_exact) {
    s_trim <- sub("0+$", "", s) # drop trailing zeros
    s_trim <- sub("\\.$", "", s_trim) # drop trailing decimal point
  } else {
    s_trim <- s
  }
  out[ok] <- ifelse(add, paste0(s, ellip), s_trim)
  out[is.na(x)] <- NA_character_
  out
}
#####################################
# function: function to format 6dp numbers with a space after the 3rd decimal
# digit
# Usage (in-line): TBI
# Usage (console): TBI
#####################################
format_spaced_6dp <- function(x) {
  # Format to 6 decimal places
  s <- sprintf("%.6f", x)
  # Insert space between the first 3 and last 3 decimal digits
  sub("(\\.[0-9]{3})([0-9]{3})$", "\\1 \\2", s)
}
#####################################
# function: format variable names in green in latex (& normal code in html)
# Usage (in-line): TBI `r format_var_name("dvisit")`
#####################################
format_var_name <- function(x) {
  if (knitr::is_latex_output()) {
    # replace all "_" with "\_" in va name
    paste0("\\greentt{", gsub("_", "\\\\_", x), "}")
  } else {
    # change to paste0("<code style='color:green'>", x, "</code>") for green
    # paste0("`", x, "`")
    paste0("<code style='color:green'>", x, "</code>")
  }
}
```

```{r params-setup, include=FALSE, echo=FALSE}
#------------------- Global parameters (used across all tasks) ----------------#
## Q1 calibration (from Excel) ----
S0    <- 392.74
sigma <- 0.269077702 
sigma <- 0.268967007212271
T_mat <- 1
N     <- 10
dt    <- T_mat / N

## CRR up/down factors ----
u <- exp(sigma * sqrt(dt))
d <- 1 / u

## Risk-neutral pricing inputs (Q3 specification) ----
r    <- 0.038
K    <- 1.05 * S0
q    <- (exp(r * dt) - d) / (u - d)
disc <- exp(-r * dt)
```

# Binomial model calibration ($u$ and $d$ estimation) {#qone}

_Note: For clarity of presentation, values in this report are displayed rounded
to a suitable level of precision. However, all underlying calculations were
performed using exact, unrounded figures to ensure accuracy._

The dataset contains 2,503 daily MSFT closing prices from 1 March 2016 to 27
February 2026. The most recent closing price is $S_0 = `r S0`$.

To calibrate a one-year binomial tree model, the historical daily closing prices
over the last ten years were accessed for Microsoft Corporation (MSFT) stock
from 1 March 2016 to 27 February 2026 using the Excel
`r format_var_name("STOCKHISTORY")` function. The daily log-return series was
calculated using

$$
r_{t} = \ln\left(\dfrac{S_t}{S_{t-1}}\right)
$$

This yields 2,502 daily log-return observations. In Excel, this was computed in
cell C10: `r format_var_name("=LN(B10/B9)")`

From the historical log-returns, the empirical daily variance was estimated
using the sample variance function (`r format_var_name("VAR.S")`) in Excel^[The
sample variance is the appropriate metric here because our 10-year historical
dataset is a sample of Microsoft's overall price history. Applying Bessel's
correction accounts for the fact that the true population mean is unknown,
yielding an unbiased estimator of the true daily variance.]:

$$
\widehat{\text{Var}}(r_t) = 0.000287313
$$

This corresponds to a daily standard deviation of
$\hat{\sigma}_{\text{daily}} = 0.016950$.

Assuming 252 trading days per year, the annualised volatility is:

$$
\hat{\sigma}_{\text{annual}} = \hat{\sigma}_{\text{daily}} \times \sqrt{252} = 0.016950 \times 15.875 = `r formatC(sigma, format = "f", digits = 6)`
$$

Using an assumption of ten time steps over one year:

$$
T = 1, \quad N = 10, \quad \Delta t = \frac{T}{N} = 0.1
$$

The up and down factors are defined using the Cox–Ross–Rubinstein (CRR)
specification:

$$
u = e^{\sigma \sqrt{\Delta t}} = e^{`r formatC(sigma, format="f", digits=6)` \times \sqrt{0.1}} = `r formatC(u, format="f", digits=5)`
\quad , \quad
d = \frac{1}{u} = \frac{1}{`r formatC(u, format="f", digits=5)`} = `r formatC(d, format="f", digits=5)`
$$

The recombining property $u \times d = 1$ holds exactly by construction.

<!-- prettier-ignore -->
::: {.answer-wrapper}
::: {.answer}
$u=`r formatC(u, format="f", digits=5)`, \quad d=`r formatC(d, format="f", digits=5)`$
:::
:::

These values imply that at each step of the binomial tree, the stock price
either increases by a factor of approximately
`r formatC(u, format="f", digits=5)` or decreases by a factor of approximately
`r formatC(d, format="f", digits=5)`.

For completeness, using the continuously compounded risk-free rate $r = `r r`$
as given in \@ref(qthr), the risk-neutral probability is:

$$
q
= \frac{e^{r\Delta t} - d}{u - d}
= \frac{e^{`r r` \times 0.1} - (`r formatC(d, format="f", digits=5)`)}{(`r formatC(u, format="f", digits=5)`) - (`r formatC(d, format="f", digits=5)`)}
= `r formatC(q, format="f", digits=5)`
$$

Although the main requirement in this part is to estimate $u$ and $d$, this
probability is useful for the construction of the binomial tree in \@ref(qtwo).

# Binomial stock price tree {#qtwo}

Figure \@ref(fig:q2-binomial-tree-plot) illustrates the full binomial tree,
showing the recombining structure and the corresponding stock prices at each
node. The tree was implemented in **R**.

```{r q2-setup, include=FALSE}
S <- matrix(0, N + 1, N + 1)
for (i in 0:N) {
  for (j in 0:i) {
    S[j + 1, i + 1] <- S0 * u^j * d^(i - j)
  }
}
```

```{r q2-binomial-tree-plot, echo=FALSE, fig.width=7, fig.height=5.5, fig.cap="CRR Binomial Stock Price Tree for MSFT over a one-year horizon with ten time steps.", fig.pos='!ht'}
par(mar = c(4, 4, 3, 1))
plot(
  NULL,
  xlim = c(-0.5, N + 1.2),
  ylim = c(min(S[S > 0]) * 0.93, max(S) * 1.04),
  xlab = "Time step (i)",
  ylab = "Stock price ($)",
  main = "CRR Binomial Stock Price Tree (MSFT, N = 10)"
)
for (i in 0:N) {
  for (j in 0:i) {
    price <- S[j + 1, i + 1]
    if (i < N) {
      lines(c(i, i + 1), c(price, S[j + 2, i + 2]), col = "steelblue", lwd = 0.7)
      lines(c(i, i + 1), c(price, S[j + 1, i + 2]), col = "steelblue", lwd = 0.7)
    }
    points(i, price, pch = 16, cex = 0.6, col = "navy")
    text(i, price, sprintf("%.2f", price), cex = 0.42, pos = 3, offset = 0.35)
  }
}
```

Using the calibrated parameters obtained in \@ref(qone), a binomial stock price tree was constructed using a discrete-time framework. The initial stock price was taken as the most recent observed value in the dataset ($S_0 = `r S0`\$$ on 27 February 2026). The binomial tree was generated over a one-year horizon with ten discrete time steps. At each step, the stock price evolves according to the CRR model, where the price either moves up by a factor $u$ or down by a factor $d$. The stock price at each node is given by:

$$
S(t,j) = S_{0} \times u^{j} \times d^{t-j}
$$

where $t$ denotes the time step and $j$ represents the number of upward movements.

The tree is recombining, meaning that different paths can lead to the same node (since $u \times d = 1$). The first-step node values are calculated as follows:

$$
S_{0} = `r S0`, \quad u = `r formatC(u, format="f", digits=5)`, \quad d = `r formatC(d, format="f", digits=5)`
$$
$$
\begin{aligned}
S(1,1) &= S_{0} \times u = `r S0` \times `r formatC(u, format="f", digits=5)` = `r formatC(S[2, 2], format="f", digits=2)` \\
S(1,0) &= S_{0} \times d = `r S0` \times `r formatC(d, format="f", digits=5)` = `r formatC(S[1, 2], format="f", digits=2)`
\end{aligned}
$$

The second-step node values are given by:
$$
\begin{aligned}
S(2,2) &= S_{0} \times u^{2} = `r formatC(S[3, 3], format="f", digits=2)` \\
S(2,1) &= S_{0} \times u \times d = `r formatC(S[2, 3], format="f", digits=2)` \\
S(2,0) &= S_{0} \times d^{2} = `r formatC(S[1, 3], format="f", digits=2)`
\end{aligned}
$$

The remaining nodes are computed using the same procedure. In **R** this was
implemented as follows:
<!-- , attr.source='.numberLines' -->
```{r q2-stock-tree, echo=TRUE}
# Building stock price matrix
# S[j+1, i+1] = S0 * u^j * d^(i-j)
S <- matrix(0, N + 1, N + 1)
for (i in 0:N) {
  for (j in 0:i) {
    S[j + 1, i + 1] <- S0 * u^j * d^(i - j)
  }
}
```

The tree has `r N + 1` terminal nodes ranging from \$`r formatC(S[1, N+1], format="f", digits=2)` (10 down moves) to \$`r formatC(S[N+1, N+1], format="f", digits=2)` (10 up moves). The initial price \$`r S0` recurs along the central path at every even step.

```{=latex}
\newpage
```

# European vanilla put option valuation {#qthr}

For a European plain vanilla put option with maturity of 1 year and a strike price equal to 105% of the initial stock price
$$
T = 1 \; \textrm{yr}, 
\quad
K = 1.05 \times S_0 = `r formatC(K, format="f", digits=3)`,
\quad
r = `r r * 100`\% \; \textrm{per annum}.
$$
where $r$ is the continuously compounded risk-free rate.

As introduced in \@ref(qone), the risk-neutral up probability is:
$$
q = \frac{e^{r\,\Delta t} - d}{u - d}
= \frac{e^{`r r` \times 0.1} - `r formatC(d, format="f", digits=5)`}{`r formatC(u, format="f", digits=5)` - `r formatC(d, format="f", digits=5)`}
= `r formatC(q, format="f", digits=5)`
$$

The terminal payoff at maturity is $V(N,j) = \max(K - S(N,j),\, 0)$. For a European option, no early exercise is permitted, so the option value at each interior node is determined purely by backward induction using the discounted risk-neutral expectation:
$$
V(i,j) = e^{-r\,\Delta t}\biggl[q \cdot \,V(i+1,\,j+1) + (1-q) \cdot V(i+1,\,j)\biggr]
$$

```{r q3-european-put, echo=TRUE, attr.source='.numberLines'}
# Q3: European Put -> backward induction 

## initialise option value matrix 
V_eu <- matrix(0, N + 1, N + 1)

## terminal payoff: max(K - S, 0) 
for (j in 0:N) {
  V_eu[j + 1, N + 1] <- max(K - S[j + 1, N + 1], 0)
}

## backward induction: discounted risk-neutral expectation 
for (i in (N - 1):0) {
  for (j in 0:i) {
    V_eu[j + 1, i + 1] <- disc *
      (q * V_eu[j + 2, i + 2] + (1 - q) * V_eu[j + 1, i + 2])
  }
}
```

Our **R** implementation first populates the terminal column with the put payoff
(lines 7-9), then works backwards through the tree (lines 12-17), computing the
discounted expected value at each node.

<!-- prettier-ignore -->
::: {.answer-wrapper}
::: {.answer}
**European put price:** $V_0 = `r formatC(V_eu[1, 1], format = "f", digits = 4)`$
:::
:::

```{r q3-tree-plot-helper, echo=FALSE}
# reusable plot function 
plot_tree <- function(V, S, title, show_early = NULL) {
  par(mar = c(4, 4, 3, 1))
  yvals <- S[S > 0]
  plot(
    NULL,
    xlim = c(-0.5, N + 1.2),
    ylim = c(min(yvals) * 0.93, max(yvals) * 1.04),
    xlab = "Time step (i)",
    ylab = "Stock price ($)",
    main = title
  )

  for (i in 0:N) {
    for (j in 0:i) {
      price <- S[j + 1, i + 1]
      val <- V[j + 1, i + 1]

      ### Draw edges to next time step
      if (i < N) {
        lines(c(i, i + 1), c(price, S[j + 2, i + 2]), col = "grey70", lwd = 0.6)
        lines(c(i, i + 1), c(price, S[j + 1, i + 2]), col = "grey70", lwd = 0.6)
      }

      ### Node colour: red = exercise, navy = positive, grey = zero
      if (!is.null(show_early) && show_early[j + 1, i + 1]) {
        col <- "red"
        cex_pt <- 1.2
      } else if (val > 0) {
        col <- "navy"
        cex_pt <- 0.8
      } else {
        col <- "grey50"
        cex_pt <- 0.5
      }
      points(i, price, pch = 16, cex = cex_pt, col = col)

      ### Node label
      lbl <- sprintf("%.2f", val)
      if (!is.null(show_early) && show_early[j + 1, i + 1]) {
        lbl <- paste0(lbl, " (E)")
      }
      text(
        i,
        price,
        lbl,
        cex = 0.42,
        pos = 3,
        offset = 0.35,
        col = if (!is.null(show_early) && show_early[j + 1, i + 1]) {
          "red"
        } else {
          "black"
        }
      )
    }
  }

  ### Legend for Bermudan
  if (!is.null(show_early)) {
    legend(
      "topleft",
      legend = c("Early exercise (E)", "Continuation", "Zero"),
      col = c("red", "navy", "grey50"),
      pch = 16,
      pt.cex = c(1.2, 0.8, 0.5),
      cex = 0.7,
      bg = "white"
    )
  }
}
```

```{r q3-european-tree-plot, echo=FALSE, fig.width=7, fig.height=8.5, fig.cap="European Put option price tree. Node labels show the option value at each node and all zero nodes are shown in grey.", fig.pos='!ht'}
## European put tree 
plot_tree(
  V_eu,
  S,
  sprintf(
    "European Put Price Tree (K=%.2f, r=%.1f%%)  |  V0 = %.4f",
    K,
    r * 100,
    V_eu[1, 1]
  )
)
```


```{=latex}
\newpage
```

# Bermudan put option valuation {#qfou}

A Bermudan put option may be exercised before maturity at pre-specified times over the lifetime of the option. In this case, the Bermudan put has the same parameters as the European put 
$$
K = `r formatC(K, format="f", digits=3)`,
\quad
r = `r r * 100`\%, 
\quad
T = 1
$$
but permits early exercise at each of the 10 tree dates.

At every node the holder takes the greater of exercising (intrinsic value $K - S$) or continuing (holding the option):
$$
V_{\text{Berm}}(i,j) 
= 
\max \biggl(K - S(i,j),\;\; e^{-r\,\Delta t}\big[q \cdot V(i+1,j+1) + (1-q) \cdot V(i+1,j)\big]\biggr)
$$

<!-- $$ -->
<!-- V_{\text{berm}}(i,j) = \max \begin{cases} -->
<!-- K-S(i,j), \\ -->
<!-- e^{-r \Delta t} \bigg[ q \cdot V(i+1, j+1) + (1 - q)\cdot V(i+1, j) \bigg] -->
<!-- \end{cases} -->
<!-- $$ -->

```{r q4-bermudan-put, echo=TRUE, attr.source='.numberLines'}
# Q4: Bermudan Put again using backward induction, now with early exercise

## initialise option value (and early exercise matrices)
V_bm  <- matrix(0, N + 1, N + 1)
early <- matrix(FALSE, N + 1, N + 1)
## terminal payoff (same as Q3 European put)
for (j in 0:N) {
  V_bm[j + 1, N + 1] <- max(K - S[j + 1, N + 1], 0)
}

## backward induction: max(intrinsic, continuation)
for (i in (N - 1):0) {
  for (j in 0:i) {
    ### continuation value: hold the option
    cont <- disc * (q * V_bm[j + 2, i + 2] + (1 - q) * V_bm[j + 1, i + 2])
    ### intrinsic value: exercise now
    exer <- K - S[j + 1, i + 1]
    ### take the maximum; flag early exercise nodes
    V_bm[j + 1, i + 1] <- max(cont, exer)
    early[j + 1, i + 1] <- (exer > cont) & (exer > 0)
  }
}
```

Note that `r format_var_name("max(cont, exer)")` on line 20
replaces the pure continuation value used in \@ref(qthr). The Boolean matrix 
`r format_var_name("early")` (line 21) records nodes where immediate exercise is
optimal (where the intrinsic value exceeds both the continuation value and
zero).

<!-- prettier-ignore -->
::: {.answer-wrapper}
::: {.answer}
**Bermudan put price:** $V_0 = `r formatC(V_bm[1, 1], format = "f", digits = 4)`$
:::
:::

```{r q4-bermudan-tree-plot, echo=FALSE, fig.width=7, fig.height=8.5, fig.cap="Bermudan Put Option Price Tree. Red nodes marked (E) indicate where early exercise is optimal and grey nodes indicate zero value. Early exercise becomes optimal in the lower portion of the tree, where the put is deep 'in-the-money' and the time value of waiting is outweighed by the immediate payoff from exercising the put.", fig.pos='!ht'}
## Bermudan put tree ----
plot_tree(
  V_bm,
  S,
  sprintf(
    "Bermudan Put Price Tree (K=%.2f, r=%.1f%%)  |  V0 = %.4f",
    K,
    r * 100,
    V_bm[1, 1]
  ),
  show_early = early
)
```

```{=latex}
\newpage
```

# Comparing the price of European and Bermudan options {#qfiv}

As shown in Table \@ref(tab:q5-comparison-table), the Bermudan put (V~0~ = `r formatC(V_bm[1, 1], format = "f", digits = 4)`) is more expensive than the European put (V~0~ = `r formatC(V_eu[1, 1], format = "f", digits = 4)`). The difference (`r formatC(V_bm[1, 1] - V_eu[1, 1], format = "f", digits = 4)`) is the early exercise premium, or the additional value the holder derives from the right to exercise before maturity.


```{r q5-comparison-table, echo=FALSE}
comparison_df <- data.frame(
  Option = c("European put", "Bermudan put", "Early exercise premium"),
  Price = c(
    formatC(V_eu[1, 1], format = "f", digits = 4),
    formatC(V_bm[1, 1], format = "f", digits = 4),
    formatC(V_bm[1, 1] - V_eu[1, 1], format = "f", digits = 4)
  )
)
kableExtra::kbl(
  comparison_df,
  col.names = c("Option", "Value (\\$)"),
  caption = "Comparison of European and Bermudan put option prices.",
  escape = FALSE,
  booktabs = TRUE,
  align = "lr"
) %>%
  kableExtra::kable_styling(full_width = FALSE, position = "center")
```


This flexibility enables the holder to exercise early when the option is sufficiently in-the-money, thereby capturing intrinsic value without being exposed to future uncertainty. This is evident in the binomial tree (Figure \@ref(fig:q4-bermudan-tree-plot)), where several nodes are identified as optimal early exercise points.

As a result, the Bermudan option includes an early exercise premium ($\approx$ `r formatC(V_bm[1, 1] - V_eu[1, 1], format = "f", digits = 4)`), representing the value of this added flexibility. Hence, the pricing relationship is consistent with theory:

$$
V_{\text{Bermudan}} \geq V_{\text{European}}
$$

# Impact of exercise frequency on Bermudan options {#qsix}

As the number of possible exercise dates increases, the value of the Bermudan option will increase (or remain unchanged), since the holder benefits from greater flexibility in choosing the optimal exercise time. With more exercise opportunities, the option can respond more effectively to favourable movements in the underlying stock price, particularly when it becomes deep in-the-money.

This reduces the likelihood of suboptimal continuation decisions and increases the overall expected payoff under the risk-neutral measure. In the limiting case, as the number of exercise dates approaches continuity, the Bermudan option converges to an American option. Therefore, the ordering of option values is:

$$
V_{\text{European}} \leq V_{\text{Bermudan}} \leq V_{\text{American}}
$$


```{=latex}
\newpage
```

<!-- # Conclusions

TBI -->

<!-- ```{=html} -->
<!-- # References -->
<!-- <div id="refs"></div> -->
<!-- ``` -->

<!-- ```{=latex} -->
<!-- \bibliography{references.bib} -->
<!-- ``` -->

```{=latex}
\newpage
\section*{Appendices}
\addcontentsline{toc}{section}{Appendices}

% --- restore default section numbering in the appendix ---
\renewcommand{\thesection}{\arabic{section}}
\renewcommand{\thesubsection}{\thesection.\arabic{subsection}}
\renewcommand{\thesubsubsection}{\thesubsection.\arabic{subsubsection}}
```

# (APPENDIX) Appendices {-}

# Supplemental Methodology {#apd-supplementary-methods}

## R implementation of binomial tree for \@ref(qtwo) {#apd:R-q2}

The binomial stock price tree was implemented in **R** using a nested loop to
populate an $(N+1) \times (N+1)$ matrix. The element 
`r format_var_name("S[j+1, i+1]")` stores the
stock price at time step $i$ with $j$ upward movements. R uses 1-based indexing,
hence the $+1$ offset. The forward induction loop iterates over all time steps
$i = 0, \ldots, N$ and, for each step, over all reachable nodes 
$j = 0, \ldots,i$, applying $S(i,j) = S_0 \cdot u^j \cdot d^{i-j}$.

## Spreadsheet implementation (Excel) {#apd:excel-impl}

This subsection outlines the structure of the accompanying Excel workbook. The calibration (estimation of $u$ and $d$) was carried out in the lightly shaded area of the provided spreadsheet...

BE note: haven't filled this out yet.


# Generative AI declaration {#apd:genai}

Generative AI (GenAI) tools were used throughout this project to assist with
project planning, methodological validation, and debugging. In practice, these
tools functioned as an additional technical resource, providing immediate
feedback, troubleshooting code, and challenging statistical assumptions.

## Tools used {-}

Google Gemini 3 Pro (Deep Research), ChatGPT 5.2, Claude (Anthropic).

## Prompts used {-}

The GenAI tools were provided with the coursework task sheet and
extracts from our R code.
A non-exhaustive list of prompts used is shown below. We opted to include at
least one example of each *type* of prompt used, rather than a list of all
prompts. For example, if we entered the prompts "How can we put two ggplots side
by side?" and "How to remove grey background on a ggplot", we would include only
the first of these in the list below, as both relate to plotting issues (and are
similar to querying with a traditional search engine).

- "...the top of both graphs in R are slightly different. (Pasting R code.) Can
  this be fixed for the plot area?"
- "Help to identify the key points here"
- "How can this error be tracked down in an Rmd file? `! Text line contains an
  invalid character. l.1 ^^@^^@...`"

## Use of GenAI outputs & subsequent revisions {-}

The GenAI tools served as a consultative resource for the group. Notably, the AI
output:

- Helped debug plotting issues and R Markdown compilation errors.
- Commented on the structure of R code snippets and the report structure.
- Assisted with converting draft content between document formats.

## Changes made {-}

All AI suggestions were critically reviewed and debated by the group before
implementation. For example:

- While the AI suggested code for complex layouts, the specific aesthetic
  choices and annotations were manually adjusted to ensure clarity.
- Where the AI proposed specific interpretations, these were cross-referenced
  with academic literature (Hull 2024, lecture notes).
- The final R code for implementation was written and debugged by the group, with
  AI snippets being treated as drafts rather than final solutions.

While GenAI was treated as a collaborator, the final analysis, statistical
interpretation, and the written content of this report were generated and
verified by the (human) group members.

